---
title: "polcon_preprocess"
author: "Erica Yoon"
date: "1/2/2019"
output: html_document
---

This markdown preprocesses polcon data: groups all polcon raw data files together to create a processed version. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse); library(data.table); library(jsonlite); library(ndjson); library(here)

# to install langcog: `install.packages("devtools"); devtools::install_github("langcog/langcog")`
library(langcog)
```

preprocess experiment 1 data.

note: experiment 1 data was hand-coded by the experimenter and therefore were already put together into a single data spreadsheet, instead of having separate file for each participant.

```{r}
d1 <- read.csv(here::here("03_analysis/data/raw_data/polcon_v1_data.csv")) %>%
  gather(trial, answer, trial1_1:trial12_2) %>%
  select(subid, order, trial, answer)

order <- read.csv(here::here("03_analysis/data/info/v1/v1_order.csv"))

d1_1 <- left_join(d1, order) %>%
  filter(order == "1") %>%
  mutate(speaker = speaker_ord1,
         request_type = request_type_ord1,
         question_type = question_type_ord1,
         item = item_ord1,
         correct = correct_ord1,
         polite_speaker = polite_speaker_ord1) %>%
  select(subid, order, trial, speaker, request_type,question_type, item, correct, polite_speaker, answer) %>%
  mutate(correct = ifelse(correct == answer, 1, 0),
         choose_polite = ifelse(polite_speaker == answer, 1, 0))

d1_2 <- left_join(d1, order) %>%
  filter(order == "2") %>%
  mutate(speaker = speaker_ord2,
         request_type = request_type_ord2,
         question_type = question_type_ord2,
         item = item_ord2,
         correct = correct_ord2,
         polite_speaker = polite_speaker_ord2) %>%
  select(subid, order, trial, speaker, request_type,question_type, item, correct, polite_speaker, answer) %>%
  mutate(correct = ifelse(correct == answer, 1, 0),
         choose_polite = ifelse(polite_speaker == answer, 1, 0))

d1 <- rbind(d1_1, d1_2)

log <- read.csv(here::here("03_analysis/data/info/v1/v1_processed_log.csv"))

d1 <- left_join(d1, log)

d_v1 <- d1
```

preprocess experiment 2 data. 

```{r}
raw.data.path <- "03_analysis/data/raw_data/v2/"
processed.data.path <- "03_analysis/data/processed_data/"

## LOOP TO READ IN FILES
all.data <- data.frame(
  subid = character(0),
  order = character(0),
  trial_num = numeric(0),
  target1side = character(0),
  target2side = character(0),
  qtype1 = character(0),
  qtype2 = character(0),
  correct1 = numeric(0),
  correct2 = numeric(0),
  date = character(0), 
  time = character(0)
)
files <- dir(here::here(raw.data.path),pattern="*.csv")


for (file.name in files) {
  print(file.name)
  
  ## these are the two functions that are most meaningful
  d <- read.csv(here::here(paste(raw.data.path,file.name,sep="")), header=FALSE)
  
  ## now here's where data get bound together
  all.data <- rbindlist(list(all.data, d))
}

d1 <- all.data %>%
  gather(var, question_type, qtype1:qtype2) %>%
  mutate(trial = trial_num) %>%
  mutate(var = substr(var, 6, 6)) %>%
  select(subid, order, var, trial, question_type)
  
d2 <- all.data %>%
  gather(var, correct, correct1:correct2) %>%
  mutate(trial = trial_num) %>%
  mutate(var = substr(var, 8, 8)) %>%
  select(subid, order, trial, var, correct)

d1 <- left_join(d1, d2) %>%
  mutate(trial = as.factor(as.character(trial))) %>%
  mutate(question_type = as.factor(ifelse(order == "2" & var == "2" & trial == "5", "comp", question_type)))

# order sheet with trial info
order <- read.csv(here::here("03_analysis/data/info/v1/v1_order.csv")) %>%
  mutate(trial = substr(trial, 6,9)) %>%
  separate(trial, c("trial", "var"))

d1_1 <- left_join(filter(d1, order == "1"), 
                  select(order, trial, var, request_type_ord1, item_ord1)) %>%
  mutate(request_type = request_type_ord1,
         item = item_ord1) %>%
  select(subid, order, trial, request_type,question_type, item, correct)

d1_2 <- left_join(filter(d1, order == "2"), 
                  select(order, trial, var, request_type_ord2, item_ord2)) %>%
  mutate(request_type = request_type_ord2,
         item = item_ord2) %>%
  select(subid, order, trial, request_type,question_type, item, correct)

d1 <- rbind(d1_1, d1_2) %>%
    mutate(valence = ifelse(question_type %in% c("polite","nice"), 
                          "positive", "negative"), 
         adjective = ifelse(question_type %in% c("nice","mean"), "easy","hard")) 

# log sheet with subj info
log <- read.csv(here::here("03_analysis/data/info/v2/subject_log.csv")) %>%
  select(-X)

d1 <- left_join(d1, log)

# make factors
d1 <- d1 %>%
  mutate(subid = as.factor(subid),
         order = as.factor(order),
         age_fct = floor(age),
         question_type = as.factor(question_type),
         gender = as.factor(gender))

# drop age other than 3 and 4, and age_fct NA
d1 <- d1 %>%
   filter(age_fct %in% c(3, 4))

d_v2 <- d1

```

preprocess experiment 3 data from nursery school sample.

```{r}
raw.data.path <- "03_analysis/data/raw_data/v3/"
processed.data.path <- "03_analysis/data/processed_data/"

## LOOP TO READ IN FILES
all.data <- data.frame(
  subid = character(0),
  order = character(0),
  trial_num = numeric(0),
  target1side = character(0),
  target2side = character(0),
  qtype1 = character(0),
  qtype2 = character(0),
  correct1 = numeric(0),
  correct2 = numeric(0),
  date = character(0), 
  time = character(0)
)
files <- dir(here::here(raw.data.path),pattern="*.csv")


for (file.name in files) {
  print(file.name)
  
  ## these are the two functions that are most meaningful
  d <- read.csv(here::here(paste(raw.data.path,file.name,sep="")), header=FALSE)

  ## now here's where data get bound together
  all.data <- rbindlist(list(all.data, d))
}

d1 <- all.data %>%
  gather(var, question_type, qtype1:qtype2) %>%
  mutate(trial = trial_num) %>%
  mutate(var = substr(var, 6, 6)) %>%
  select(subid, order, var, trial, question_type)
  
d2 <- all.data %>%
  gather(var, correct, correct1:correct2) %>%
  mutate(trial = trial_num) %>%
  mutate(var = substr(var, 8, 8)) %>%
  select(subid, order, trial, var, correct)

d1 <- left_join(d1, d2) %>%
  mutate(trial = as.factor(as.character(trial))) %>%
  mutate(question_type = as.factor(ifelse(order == "2" & var == "2" & trial == "5", "comp", question_type)))

# order sheet with trial info
order <- read.csv(here::here("03_analysis/data/info/v1/v1_order.csv")) %>%
  mutate(trial = substr(trial, 6,9)) %>%
  separate(trial, c("trial", "var"))

d1_1 <- left_join(filter(d1, order == "1"), 
                  select(order, trial, var, request_type_ord1, item_ord1)) %>%
  mutate(request_type = request_type_ord1,
         item = item_ord1) %>%
  select(subid, order, trial, request_type,question_type, item, correct)

d1_2 <- left_join(filter(d1, order == "2"), 
                  select(order, trial, var, request_type_ord2, item_ord2)) %>%
  mutate(request_type = request_type_ord2,
         item = item_ord2) %>%
  select(subid, order, trial, request_type,question_type, item, correct)



d1 <- rbind(d1_1, d1_2)

# log sheet with subj info
log <- read.csv(here::here("03_analysis/data/info/v3/subject_log_v3.csv"))

d1 <- left_join(d1, log)

# filter by exclusion criteria
d1 <- d1 %>% 
  filter(keepdrop != "drop", age >= 3 | age <= 4)
  
# make factors
d1 <- d1 %>%
  mutate(subid = as.factor(subid),
         order = as.factor(order),
         question_type = as.factor(question_type),
         gender = as.factor(gender))

# how many before filtering by practice?
subj <- d1 %>%
    distinct(subid, .keep_all=TRUE) %>%
    mutate(age_fct = floor(age)) %>%
    group_by(age_fct) %>%
    summarise(count = n())
subj

# change question_type
# v3 had just polite vs. rude, but this was not reflected in the logging system 

d1 <- d1 %>%
  mutate(question_type = fct_recode(question_type,
                                    "polite" = "nice",
                                    "rude" = "mean")) %>%
  # make age grouping
  mutate(age_fct = floor(age)) %>%
  filter(question_type != "comp") %>%
  select(-keepdrop, -initials, -dot, -dob)

d_v3 <- d1

```

preprocess expt3 data from online sample.

```{r}
# create function for calculating age.
calc_age <- function(birthDate, refDate = Sys.Date()) {

    require(lubridate)

    period <- as.period(interval(birthDate, refDate),
                        unit = "day")

    round(period$day/365, 2)

}

data_path <- "03_analysis/data/raw_data/v3_lookit"
file <- "polcon_lookit_raw_data.csv"
# file <- "test.csv"
d_raw <- read_csv(here::here(data_path, file))

ord <- read_csv(here::here("03_analysis/data/info/v3/polcon_order.csv"))

d_tidy <- d_raw %>%
  # filter(response_completed == "True") %>%
  # filter(participant_id == "3434" | participant_id == "3432") %>%
  filter(!grepl("test", child_additional_information, ignore.case=TRUE)) %>%
  # mutate(order = case_when(
  #   grepl("'conditionNum': 0", substr(response_conditions, 1, 108)) ~ "1",
  #   grepl("'conditionNum': 1", substr(response_conditions, 1, 108)) ~"2"
  #   )) %>%
  mutate(response_conditions = gsub("\'","\"",response_conditions)) %>%
  mutate(response_exp_data = gsub("\'","\"",response_exp_data),
         response_exp_data = gsub("None","\"None\"",response_exp_data),
         response_exp_data = gsub("True","\"True\"",response_exp_data),
         response_exp_data = gsub("False","\"False\"",response_exp_data),
         response_exp_data = gsub("\"s ","\'s ",response_exp_data),
         response_exp_data = gsub("\"rude\"","\\'rude\\'",response_exp_data),
         response_exp_data = gsub("I\"m ","I\'m ",response_exp_data),
         response_exp_data = gsub("I\"d ","I\'d ",response_exp_data),
         response_exp_data = gsub("they\"d ","they\'d ",response_exp_data),
         response_exp_data = gsub("n\"t ","n\'t ",response_exp_data),
         response_exp_data = gsub("\"please\"","\'please\'",response_exp_data),
         response_exp_data = gsub("\"can ","\'can ",response_exp_data),
         response_exp_data = gsub("please\".\"","please\'.\"",response_exp_data),
         response_exp_data = gsub("\"\"None\"\"","\"None\"", response_exp_data),
         response_exp_data = gsub("\"Next\"","\\'Next\\'",response_exp_data)

         ) %>%
  select(response_id, response_conditions, child_birthday, child_gender, response_exp_data) %>%
  rename(subid = response_id) %>%
  mutate(
    # order 1 = cond 0, order 2 = cond 1
    order_practice =
      jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_conditions)))$`5-practice-trial`$conditionNum,
    order_test = 
      jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_conditions)))$`6-nested-trials`$conditionNum,

    dot = 
      jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`0-0-video-config`$eventTimings,
      
    practice1 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`9-9-practice-trial`$currentlyHighlighted) %>%
  filter(!is.na(practice1)) %>%
  mutate(
    practice2 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`10-10-practice-trial`$currentlyHighlighted,
    test1 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`16-16-nested-trials`$currentlyHighlighted,
    test2 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`22-22-nested-trials`$currentlyHighlighted,
    test3 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`28-28-nested-trials`$currentlyHighlighted,
    test4 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`34-34-nested-trials`$currentlyHighlighted,
    test5 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`40-40-nested-trials`$currentlyHighlighted,
    test6 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`46-46-nested-trials`$currentlyHighlighted,
    test7 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`52-52-nested-trials`$currentlyHighlighted,
    test8 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`58-58-nested-trials`$currentlyHighlighted,
    test9 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`64-64-nested-trials`$currentlyHighlighted,
    test10 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`70-70-nested-trials`$currentlyHighlighted,
    test11 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`76-76-nested-trials`$currentlyHighlighted,
    test12 = jsonlite::stream_in(textConnection(gsub("\\n", "", .$response_exp_data)))$`82-82-nested-trials`$currentlyHighlighted
         ) %>%
    rowwise() %>%
    mutate(dot = as.Date(substr(dot$timestamp[1], 1, 10))) %>%
    filter(dot > "2018-04-05") %>%
    mutate(age = calc_age(child_birthday, dot))
```

```{r}
subj_v3_online_too_few_trials <- d_tidy %>% 
  select(subid, test1:test12) %>%
  gather(key=trial_num, value=answer, -subid) %>%
  filter(!is.na(answer)) %>%
  distinct(subid, trial_num, .keep_all=T) %>%
  count(subid) %>%
  filter(n < 6) %>%
  mutate(reason_for_drop = "too few trials") %>%
  select(subid, reason_for_drop)

subj_v3_online_failed_practice <- d_tidy %>%
  filter((order_practice == 0 & (practice1 != "speaker1" | practice2 != "speaker2")) |          (order_practice == 1 & (practice1 != "speaker2" | practice2 != "speaker1"))) %>%
  mutate(reason_for_drop = "failed practice") %>%
  select(subid, reason_for_drop)

subj_excl <- rbind(subj_v3_online_too_few_trials, subj_v3_online_failed_practice)

write.csv(subj_excl,here::here(paste("03_analysis/data/processed_data/v3_drop.csv",sep="")),
           row.names=FALSE)
```


```{r}
d <- d_tidy %>%
  filter(!is.na(test6)) %>%
  filter((order_practice == 0 & practice1 == "speaker1" & practice2 == "speaker2") |
           (order_practice == 1 & practice1 == "speaker2" & practice2 == "speaker1")) %>%
  select(-order_practice, -practice1, -practice2) %>%
  gather(trial_num, choice, test1:test12) %>%
  mutate(trial_num = as.numeric(as.character(substr(trial_num, 5, 6)))) %>%
  left_join(., ord) %>%
  mutate(correct = as.numeric(as.character(case_when(
    choice == correct_answer ~ "1",
    choice != correct_answer ~ "0",    
    TRUE ~ "NA"
  ))))

d_v3_lookit <- d

```



bind all data together.

```{r}
d_v1 <- d_v1 %>%
  mutate(expt = "v1", 
         trial = substr(trial, 6, 6))

d_v2 <- d_v2 %>%
  mutate(expt = "v2")

d_v3 <- d_v3 %>%
  mutate(expt = "v3")

d_v3_lookit <- d_v3_lookit %>%
  mutate(expt = "v3_lookit") %>%
  mutate(order = case_when(
    order_test == "0" ~ "1",
    order_test == "1" ~ "2",
    TRUE ~ "NA"
    )) %>%
  rename(trial = trial_num, question_type = question)

d <- rbind(
  d_v1 %>% rename(gender = sex) %>% select(subid, expt, age, gender, order, trial, request_type, question_type, item, correct) %>%
    mutate(subid = as.character(subid)),
  d_v2 %>% 
    select(subid, expt, age, gender, order, trial, request_type, question_type, item, correct) %>%
    mutate(subid = as.character(subid)),
  d_v3 %>% select(subid, expt, age, gender, order, trial, request_type, question_type, item, correct) %>%
    mutate(subid = as.character(subid)),
  d_v3_lookit %>% 
    mutate(gender = fct_recode(child_gender, "M" = "m", "F" = "f", "NA" = "na")) %>%
    select(subid, expt, age, gender, order, trial, request_type, question_type, item, correct)
) %>%
  mutate_if(is.character, as.factor) %>%
  mutate(subid = as.character(subid)) %>%
  mutate(age_fct = floor(age)) %>%
# drop age other than 3 and 4, and age_fct NA
   filter(age_fct %in% c(2, 3, 4)) %>%
  mutate(question_type = fct_recode(question_type,
    "comp" = "compliance",
    "play" = "play_with"
  )) %>%
  mutate(correct = as.numeric(as.character(correct))) %>%
  mutate(question_type = fct_relevel(question_type, 
                                     "polite", "rude", "nice", "mean"),
         request_type = fct_recode(request_type,
                                   "can you~" = "idq",
                                   "can you please~" = "please_idq"),
         request_type = fct_relevel(request_type, "please")) %>%
  mutate(subid = as.factor(subid)) %>%
  filter(gender != "NA")

write.csv(d,here::here(paste("03_analysis/data/processed_data/polcon_all_data.csv",sep="")),
           row.names=FALSE)

```

