---
title: "Children's understanding of simple polite markers"
bibliography: polcon_cogsci2019.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf } \\ \texttt{} \\  \\}

abstract: 
    ""
    
keywords:
    "Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term."
    
output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)

library(here)
library(langcog)
```

```{r load_data}
d <- read_csv(here::here("03_analysis/data/processed_data/polcon_all_data.csv")) %>%
  filter(!is.na(gender) & gender != "na")
```


# Introduction

As adults, we use polite speech all the time.

Children produce polite speech early on.

Less work has looked at children's comprehension of polite speech.

For example, @nippold1982 looked at...

In this current work, we seek to test what 2- to 4-year-old children understand about polite speech. Specifically, we ask: (1) whether children can reason about which speaker is being more "polite" or "rude" (or "nice" or "mean"); (2) whether they understand possible social consequences of being polite or impolite; and (3) how this reasoning may change across development.

# Experiment 1

## Methods

### Participants

```{r}
# exclusion criteria

# exclusion criteria applied to expt 1.
subj_v1_too_few_trials <- d %>% 
  filter(expt == "v1", trial == "6", is.na(correct)) %>% 
  distinct(subid)

subj_v1_too_few_trials_count <- subj_v1_too_few_trials %>% 
  count(subid)
# 
# d <- d %>%
#   filter(!subid %in% subj_v1_too_few_trials$subid)

# exclusion criteria applied to expt 2.
subj_v2_failed_practice <- filter(d, (expt == "v2" & trial == "0" & correct == "0")) %>% 
  distinct(subid)

subj_v2_failed_practice_count <- subj_v2_failed_practice %>%
  count(.)
# 
# d <- d %>%
#   filter(!subid %in% subj_v2_failed_practice$subid)

subj_v2_too_few_trials <- d %>% 
  filter(expt == "v2") %>% 
  count(subid) %>%
  filter(n < 6)

subj_v2_too_few_trials_count <- subj_v2_too_few_trials %>% 
  count(.)
# 
# d <- d %>%
#   filter(!subid %in% subj_v2_too_few_trials$subid)

# exclusion criteria applied to expt 3 in-person sample.
subj_v3_failed_practice <- filter(d, (expt == "v3" & trial == "0" & correct == "0")) %>% 
  distinct(subid)

subj_v3_failed_practice_count <- subj_v3_failed_practice %>%
  count(.)
# 
# d <- d %>%
#   filter(!subid %in% subj_v3_failed_practice$subid) %>%
#   filter(!is.na(correct))

subj_v3_too_few_trials <- d %>% 
  filter(expt == "v3") %>% 
  distinct(subid, trial, .keep_all=T) %>%
  count(subid) %>%
  filter(n < 6)

subj_v3_too_few_trials_count <- subj_v3_too_few_trials %>% 
  count(.)

d <- d %>% 
  filter(!subid %in% subj_v1_too_few_trials$subid) %>%
  filter(!subid %in% subj_v2_failed_practice$subid) %>%
  filter(!subid %in% subj_v2_too_few_trials$subid) %>%
  filter(!subid %in% subj_v3_failed_practice$subid) %>%
  filter(!subid %in% subj_v3_too_few_trials$subid)
```

```{r}
# exclusion from online sample was computed separately before preprocessing
v3_online_excl <- read_csv(here::here("03_analysis/data/processed_data/v3_drop.csv")) %>%
  count(reason_for_drop)

v3_online_excl_few_trials <- v3_online_excl %>% 
  filter(reason_for_drop == "too few trials")

v3_online_excl_failed_practice <- v3_online_excl %>% 
  filter(reason_for_drop == "failed practice")
```

```{r}
subj <- d %>% 
  count(expt, age_fct, gender, subid) %>%
  group_by(expt, age_fct, gender) %>%
  summarise(n=n())

ms_age <- d %>% 
  group_by(expt, age_fct) %>%
  summarise(m = mean(age, na.rm = T),
            sd_age = sd(age, na.rm = T)) %>% 
  mutate_all(round, 2)

subj_v1_3y <- subj %>% filter(expt == "v1", age_fct == "3")
subj_v1_4y <- subj %>% filter(expt == "v1", age_fct == "4")

```

were recruited from a local preschool. 
An additional 3 children were tested but excluded due to failure on the practice questions (2) or completion of fewer than half of the test trials (`r subj_v1_too_few_trials_count$n`).

### Stimuli

### Procedure

## Results and Discussion

# Experiment 2

## Methods

### Participants

### Stimuli

### Procedure

## Results and Discussion

# Experiment 3

## Methods

### Participants

### Stimuli

### Procedure

## Results and Discussion

# General Discussion

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
